---
title: "R Presentation"
subtitle: "R Language Basics <br/>Using &#x2694;&#xFE0F; xaringan +<br/>&#x1F60E; xaringanthemer"  
author: "Aswin Chidambaram Balakrishnan"
date: '2022-10-01'
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: true

---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width=9, fig.height=3.5, fig.retina=3,
  out.width = "100%",
  cache = FALSE,
  echo = TRUE,
  message = FALSE, 
  warning = FALSE,
  hiline = TRUE
)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_duo_accent(
  primary_color = "#12568F",
  secondary_color = "#FF961C",
  inverse_header_color = "#FFFFFF"
)
```

### R Language

R is a programming language for **statistical computing** and graphics supported by the __R Core Team__ and the __R Foundation for Statistical Computing__.

According to user surveys and studies of scholarly literature databases, R is one of the most commonly used programming languages used in __data mining__.

It is written primarily in **C, Fortran, and R** itself (partially self-hosting).

R has a **command-line interface** and other multiple third party **graphical user interfaces**.

R is a **intepreted programming language** and it also supports **procedural programming with functions** and, for some functions, **object-oriented programming with generic functions**.

R has stronger object-oriented programming facilities than most statistical computing languages.

R's extensible object system includes objects for (among others): **regression models, time-series and geo-spatial coordinates**.

---
name: R Basics

### R Basics

#### Variable assignment in R

Variable assignment in R is done using \<- operator with variable at the left side and its corresponding value at the right side of this operator.

```{r echo=TRUE}
a <- 1
a
```

#### Functions in R

Function is used especially for the re-usability of the code and this can be called any number of times. The syntax is below:

```{r echo=TRUE}
hello <- function() {
  print("Hello!!!")
}
hello()
```

---

### Inbuilt Functions

#### Math Functions

abs(x) ---> Returns the absolute value of the input x.

```{r echo=TRUE}
x <- 10
print(abs(x))
```

sqrt(x) ---> Returns the square root value of the input x.

```{r echo=TRUE}
x <- 25
print(sqrt(x))
```

log(x) ---> Returns the natural logarithm of the input x.

```{r echo=TRUE}
x <- 5
print(log(x))
```
---

### Inbuilt Functions

#### String Functions

substr(x, start=n1, stop=n2) ---> Extract substring in a character vector.

```{r echo=TRUE}
x <- "hello!!!!"
print(substr(x, 2, 5))
```

paste(..., sep="") ---> Concatenates strings after using sep string to separate them.

```{r echo=TRUE}
x <- "hello!!!!"
paste('hi','hello',1,2,3)
```

toupper(x) ---> Converts the string into upper case.

```{r echo=TRUE}
x <- "hello!!!!"
print(toupper(x))
```